---
title: "Titanic Analysis"
output: html_document
author: "Luis Fernando García Acosta"
---

## Titanic Data Analysis
<br/>
In this pratice we have a Titanic dataset so the goal is to get some information about what happened that day or some data that we can only get from our data. We are going to start with basic data. In this pratice I'm going to use some statistics techniques or knowledges that we can aply in here to see what conclution we can get. 
<br/><br/>



## Basic Data
First I'm going to show you a general summary of what our dataset looks like.
```{r,echo=FALSE}
titanic_data <- read.csv("titanic.csv",header = TRUE)
summary(titanic_data)
```
<br/>
The first thing we can notice is that passengers have an ID so we can say that the total number of passengers is:
```{r,echo=FALSE}
nrow(titanic_data)
```
<br/>
We can get other clear information from this data we can see that in the Titanic were **314 females** and **577 males**
<br/>
<br/>
Also we can see that we have the passanger's age the minimun age was about ** 1 year old ** and the maximum age was about **80 years old**
<br/>
<br/>
And the last fact about the data before continuing with the plots, the minimu ticket fare costed **0.0 British Pounds** and the most expensive ticket costed **512 British Pounds**
<br/>
</br>
</br>
We can see that **168 persons** embarked in Cherbourg</br>
**77 persons** embarked in Queenstown</br>
**644 persons** embarked in Southampton </br>
</br>
</br>


## Data Analysis Session
<br/>
<br/>

## Age
  
The first thing we are going to analize the age of the passengers
</br>
```{r,echo=FALSE}
    summary(titanic_data$Age)
```

We can see that the minimum age was .042 years old exactly, and the maximum age was 80 years old. We can also see that the median age was 28 and the average was between 28 an 30 years old
<br/>

We can see the box plot how ages are distributed
```{r, echo=FALSE}
  boxplot(titanic_data$Age)
```

<br/> 
And the histogram shows that the most of the ages were between 20 and 30 years old, really young people

```{r,echo=FALSE}
  hist(titanic_data$Age)
```


## Class

The second thing we are going to analise is the class of the passengers
</br>
```{r,echo=FALSE}
  summary(titanic_data$Pclass)
```
First we have to understand that there were 3 classes, first class,second and third. <br/>
<br/>
<br/>
The most of the people were in third class

```{r,echo=FALSE}
getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
getmode(titanic_data$Pclass)
```
As we can see in the box plot<br/>
```{r,echo=FALSE}
boxplot(titanic_data$Pclass)
```
<br/>
<br/>
And we can get a estimation how many people were in which class with the next histogram
```{r,echo=FALSE}
hist(titanic_data$Pclass)
```

The finally conclusion I can get from this data is that in the disaster the most of the corpses did not have the chance of getting rescued from the sea, the first class passengers(corpses) had the the major priority of been rescued from the sea and we can see tha the first class passengers were like about 220-230. 

## Survivors

We can get from the data  about how many people survived we can see the next histogram of suvivors, 0 means no survivor and 1 means survivor.
```{r,echo=FALSE}
  hist(titanic_data$Survived)
```

As we can see in the histogram the most people did not survive. We can see it if we get the mode of the dataset in this column.
```{r,echo=FALSE}
  getmode(titanic_data$Survived)
```






